Emittance preservation of an electron beam in a loaded quasi-linear plasma wakefield

### 2017-09-05 Tex/Paper.tex
Last additions after last meeting w/Erik

A major challenge with plasma wakefield accelerators is, however, to produce an accelerated beam while keeping growth in
energy spread and emittance as low as possible. In the well described linear case where the beam density $n_{b}$ is much
smaller than the plasma density $n_{0}$, the non-linear transverse wakefields cause emittance growth of the accelerated
beam. A finite length beam will also see a transversely varying accelerating field causing increasing energy spread
\cite{katsouleas:1987}. In the non-linear regime, where $n_{b} > n_{0}$, the drive beam sees a constant transverse field
due to a bubble forming in the plasma behind the drive beam. The bubble is formed by the transverse oscillations of the
plasma electrons in the wake of the drive beam, forming a sheet around an evacuated area. The ions, assumed stationary,
form an ion channel creating a focusing force that varies linearly with radius. This
produces an axially symmetric focusing force \cite{lu:2006-1, lu:2006}.
%~ This effect also avoids the issue of increasing energy spread due to transverse beam loading. Increase in energy spread
%~ is still dependent on beam loading $E_{z}$.
%~ Ideal beam loading can be achieved by specially shaped beams that optimally flattens the field
%~ \cite{katsouleas:1987, tzoufras:2009}, but this is difficult experimentally and is not a topic explored in this paper.

@@

The energy carried by electron drive bunches used in previous experiments were typically small, on the order of
$100\unit{J}$, and the propagation length typically $<1\unit{m}$ \cite{blumenfeld:2007,caldwell:2009}. De-phasing of the
witness beam relative to the drive beam, caused by loss of energy in the drive beam and gain of energy in the witness
beam, limits the effective length of a plasma accelerator stage. Proton drive beams available at CERN carry
significantly more energy, $19\unit{kJ}$ for the SPS beam \cite{gschwendtner:2016}, allowing for longer accelerator
stages. However, the SPS beam is orders of magnitude longer than the plasma wavelengths needed for such applications.
This issue is resolved by letting the proton beam undergo self-modulation before injecting the witness beam into one of
the buckets in the modulated structure. This self-modulation instability is caused by the transverse fields generated by
the beam acting upon the beam itself, causing regions of the beam to rapidly defocus \cite{kumar:2010}. The modulation
frequency is close to that of the plasma, and leaves a train of proton bunches along the beam axis with a surrounding
halo.

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQuality}
    \caption{\label{Fig:BeamQ} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial beam charge (blue), and relative energy spread (red), after $4\unit{m}$ of plasma and with
        an initial emittance $\epsilon_{N,0}=2\unit{\mu m}$. These are shown for four different $\sigma_{z}$ from
        $40\unit{\mu m}$ to $100\unit{\mu m}$. The detailed studies presented in beam loading section corrspond to the
        square marked lines at $100\unit{pC}$.}
\end{figure}

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityEmittance}
    \caption{\label{Fig:BeamQEmit} }
\end{figure}

[Discussion of new emittance scan figure(s) to be added.]

>>>>

A major challenge with plasma wakefield accelerators is, however, to produce an accelerated beam while keeping growth in
energy spread and emittance as low as possible. In the well described linear case where the beam density $n_{b}$ is much
smaller than the plasma density $n_{0}$, the non-linear transverse wakefields cause emittance growth of the accelerated
beam. A finite length beam will also see a transversely varying accelerating field causing increasing energy spread
\cite{katsouleas:1987}. In the non-linear regime, where $n_{b} > n_{0}$, a bubble is formed by the transverse
oscillations of the plasma electrons forming a sheet around an evacuated area leaving only ions. The ions, assumed
stationary, form an ion channel creating a focusing force that varies linearly with radius. This produces an axially
symmetric focusing force \cite{lu:2006-1, lu:2006}.
%~ This effect also avoids the issue of increasing energy spread due to transverse beam loading. Increase in energy spread
%~ is still dependent on beam loading $E_{z}$.
%~ Ideal beam loading can be achieved by specially shaped beams that optimally flattens the field
%~ \cite{katsouleas:1987, tzoufras:2009}, but this is difficult experimentally and is not a topic explored in this paper.

@@

The energy carried by electron drive bunches used in previous experiments were typically small, on the order of
$100\unit{J}$, and the propagation length typically $<1\unit{m}$ \cite{blumenfeld:2007,caldwell:2009}. The energy of a
high-charge electron beam accelerated to $1\unit{TeV}$, similar to the beam in the international linear collider with
$1\mexp{10}$ electrons, is $1.6\unit{kJ}$. By using electron beams or lasers as drivers a large number of plasma stages
is required. How to efficiently stage plasma accelerators is challenging without reducing the effective gradient and
spoiling the beam quality \cite{steinke:2016,lindstrom:2016}. Proton drive beams available at CERN carry
significantly more energy, $19\unit{kJ}$ for the SPS beam \cite{gschwendtner:2016}, allowing for longer accelerator
stages. However, the SPS beam is orders of magnitude longer than the plasma wavelengths needed for such applications.
This issue is resolved by letting the proton beam undergo self-modulation before injecting the witness beam into one of
the buckets in the modulated structure. This self-modulation instability is caused by the transverse fields generated by
the beam acting upon the beam itself, causing regions of the beam to rapidly defocus \cite{kumar:2010}. The modulation
frequency is close to that of the plasma, and leaves a train of proton bunches along the beam axis with a surrounding
halo.
%~ De-phasing of the
%~ witness beam relative to the drive beam, caused by loss of energy in the drive beam and gain of energy in the witness
%~ beam, limits the effective length of a plasma accelerator stage.

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQuality}
    \caption{\label{Fig:BeamQ} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial beam charge (blue), and relative energy spread of the accepted charge (red), after
        $4\unit{m}$ of plasma and with an initial emittance $\epsilon_{N,0}=2\unit{\mu m}$. These are shown for four
        different $\sigma_{z}$ from $40\unit{\mu m}$ to $100\unit{\mu m}$. The detailed studies presented in beam
        loading section corrspond to the square marked lines at $100\unit{pC}$.}
\end{figure}

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityEmittance}
    \caption{\label{Fig:BeamQEmit} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial emittance (blue), and relative energy spread of the accepted charge (red), after $4\unit{m}$
        of plasma.}
\end{figure}

Figure \ref{Fig:BeamQEmit} shows how the growth in emittance and energy spread varies with initial electron beam
emittance. The smaller the initial emittance is, the better the emittance is preserved. There are two effects that lead
to growth for high emittance beams; the transverse beam size may increase beyond the size of the bubble, or the beam
density may be reduced so much that the plasma electrons are no longer fully evacuated from the ion column. Emittance
values higher than a few micrometers lead to a significant increase in both emittance and energy spread. In these
simulations the focusing of the beam to the plasma remains matched for each emittance value.
